---
title: "ggplot exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height=4, fig.width=6)
library(dplyr)
library(ggplot2)

nhanes = read.csv("height_gender.csv")
```

# Bivariate data

## Scatter plots

Scatter plots are easy to make with ggplot!

```{r petal_points}
ggplot(iris, aes(Petal.Length, Petal.Width)) +
  geom_point()
```

You can add additional aesthetics including:

* color
* shape
* size

```{r petal_points_by_species}
ggplot(iris, aes(Petal.Length, Petal.Width, color=Species, shape=Species)) +
  geom_point()
```

*Challenge*: from the NHANES data set, make a scatter plot of weight vs height,
where color indicates sex.

```{r nhanes_height_weight_by_sex}

```

## 2D density plots

Density plots are just as easy to make:

```{r petal_density}
ggplot(iris, aes(Petal.Length, Petal.Width)) +
  geom_density2d()
```


*Challenge*: from the NHANES data set, make a density plot of weight vs height,
where color indicates sex.

```{r nhanes_height_weight_density}

```

# Semicategorical bivariate data

Boxplots are a great way to visualize a continuous variable against a
categorical variable:

```{r iris_box}
ggplot(iris, aes(Species, Petal.Length)) +
  geom_boxplot()
```

You can also overlay multiple geoms on a single plot:

```{r iris_violin}
ggplot(iris, aes(Species, Petal.Length)) +
  geom_violin() +
  geom_point(position=position_jitter(width=0.2, height=0), alpha=0.5)
```

*Challenge*: Make a violin plot of height vs sex.

```{r nhanes_box}

```

# Continuous univariate data

Histograms are my favorite kind of plot! They are very hard to make in Excel.
They are very easy to make with ggplot.

```{r petal_length_density}
ggplot(iris, aes(Petal.Length)) +
  geom_histogram(bins=10) +
  facet_grid(Species~.)
```

You can also connect the tops of the histogram boxes:

```{r petal_length_freqpoly}
ggplot(iris, aes(Petal.Length, color=Species)) +
  geom_freqpoly(binwidth=0.5)
```

Density plots (the continuous analogue of histograms) are also nice to overlay.

```{r petal_length_density}
ggplot(iris, aes(Petal.Length, color=Species)) +
  geom_density()
```

*Challenge*: Make histograms of height for men and women born in the US and overseas.

```{r nhanes_histo}

```

# Bar plots

Bar charts are actually relatively difficult to make with ggplot2. You shouldn't
be making them anyway! But you can.

```{r iris_bar_plot}

# You can use ggplot's stat_summary to help you compute the mean of observations
# for a particular combination of aesthetics.

g = ggplot(iris, aes(Species, Petal.Length)) +
  stat_summary(fun.y=mean, geom="bar")
print(g)

# Or you can use dplyr to compute it in advance and use stat_identity. This
# happens to be much faster for large data sets.

g = iris %>%
  group_by(Species) %>%
  summarize(Petal.Length.mean=mean(Petal.Length)) %>%
  ggplot(aes(Species, Petal.Length.mean)) +
    geom_bar(stat="identity", width=0.5, fill="blue")
print(g)

```


Challenge: use a bar plot to compare average height for men and women in the
NHANES data set.

```{r nhanes_bar_plot}

```
